### 2020-05-10 chapter1.Rmd
edited scratchpad chapter1 and made more organic project.Rmd

```{r cleanup, echo=FALSE}
out_domain <- (seis$min < domain[1] | seis$max > domain[2])
invalid_values <- (seis$min == -1 & seis$max == 1) | (seis$min == 0 & seis$max == 0) | out_domain
"Removing " %s+% sum(out_domain) %s+% " values out of domain, " %s+% (sum(invalid_values) - sum(out_domain)) %s+% " explicit NA"
seis <-
  seis %>%
  mutate(min=replace(min, invalid_values, NA), max=replace(max, invalid_values, NA)) %>%
  drop_na()
# %>% mutate(belgian_second=with_tz(utc_second, "Europe/Brussels"))
remove(out_domain, invalid_values) # free ~350 Mb from RAM
seis.present <- length(seis$utc_second)
seis.total <- max(as.numeric(seis$utc_second)) - min(as.numeric(seis$utc_second))
seis.missing <- seis.total - seis.present
seis.present %s+% " present values, " %s+% seis.missing %s+% " missing points (" %s+% print_dec(seis.present / seis.total * 100, 2) %s+% "% / " %s+% print_dec(seis.missing / seis.total * 100, 2) %s+% "%)"

@@

@@

fitted <- fitma.list[order(aic_vals$AIC)][[2]]
tsdisplay(residuals(fitted))

@@

```{r fourier}
seis.h.start <- with_tz(as_datetime('2018-11-01T00:00:00'), "Europe/Brussels")
seis.h.end <- with_tz(as_datetime('2020-03-01T00:00:00'), "Europe/Brussels")
seis.h.msts <- build_ts(seis.h, periods = c(24, 168), start = seis.h.start, end = seis.h.end)
fitted <- auto.arima(seis.h.msts, D=0, max.P=0, max.Q=0, xreg=fourier(seis.h.msts, c(3, 4)))
plot(fitted)
plot(residuals(fitted))
tsdisplay(residuals(fitted))
res.xts <- ts_zoo(residuals(fitted))
res.msts <- msts(res.xts, c(24, 168))
res.aic_vals_temp <- NULL
res.aic_vals <- NULL
for (i in 1:5) {
  for (j in 1:5) {
    #for (k in 1:5) {
      res.xreg.d <- fourier(res.msts, c(i, j)  )  # daily  seasonality
      res.fitma <- auto.arima(res.msts, D=0, max.P=0, max.Q=0, xreg=res.xreg.d)
      res.aic_vals_temp <- cbind(i, j, res.fitma$aic)  #,k
      res.aic_vals <- rbind(res.aic_vals, res.aic_vals_temp)
      print("(" %s+% i %s+% ", " %s+% j %s+% ") = " %s+% res.fitma$aic)
    #}
  }
}
remove(res.aic_vals_temp, res.xreg.d, res.fitma)
colnames(res.aic_vals) <- c("Fourier24", "Fourier168", "AIC")
res.aic_vals <- data.frame(res.aic_vals)
res.min_aic <- min(res.aic_vals$AIC)
res.minvals <- res.aic_vals[which(res.aic_vals$AIC == res.min_aic),]
res.p <- ggplot(res.aic_vals, aes(Fourier24, Fourier168)) + geom_tile(aes(fill = AIC), colour = "white") + scale_fill_gradient(low = "white", high = "steelblue")
plot(res.p)
res.minvals
res.fitted <- auto.arima(res.msts, D=0, max.P=0, max.Q=0, xreg=fourier(res.msts, c(res.minvals$Fourier24, res.minvals$Fourier168)))
plot(res.fitted)
tsdisplay(residuals(fitted))
tsdisplay(residuals(res.fitted))

```{r periodgram}
library(spec)
#pgram <- spec.pgram(build_ts(seis.h, start=as_datetime('2018-11-01T00:00:00'), end=as_datetime('2020-03-01T00:00:00')), main="periodgram", detrend=F, demean=T, na.action=na.aggregate)
pgram <- spec.pgram(build_ts(seis.h, start=as_datetime('2019-05-03T00:00:00'), end=as_datetime('2020-03-01T00:00:00')), main="periodgram", detrend=F, demean=T, na.action=na.aggregate)

topfreq <- pgram$freq[order(pgram$spec, decreasing = T)][1:100]
topseas <- 1/topfreq
topseas.h <- topseas/3600

top95pc <- pgram$spec[pgram$spec > (mean(pgram$spec) + sd(pgram$spec)*1)]
top95freq <- pgram$freq[pgram$spec > (mean(pgram$spec) + sd(pgram$spec)*1)][order(top95pc, decreasing = T)]
top95pc <- pgram$spec[pgram$spec > (mean(pgram$spec) + sd(pgram$spec)*1)][order(top95pc, decreasing = T)]
top95seas.s <- 1/top95freq
top95seas.h <- top95seas.s/3600
top95pc/sd(pgram$spec)
top95seas.h

ylims <- 10^par("usr")[3:4]
axis(side=3, at=top95freq, labels=as.integer(top95seas.h), padj = 1, las=2, cex.axis=0.8)
abline(h=mean(pgram$spec) + sd(pgram$spec)*c(.5,1:3), lty="dotted", col=c("orange", "darkgreen", "blue", "red"))
for (i in 1:length(top95freq)) {
  color <- ifelse(top95pc[i] > (mean(pgram$spec) + sd(pgram$spec)*3), "red", ifelse(top95pc[i] > (mean(pgram$spec) + sd(pgram$spec)*2), "blue", ifelse(top95pc[i] > (mean(pgram$spec) + sd(pgram$spec)*1), "darkgreen", "orange")))
  lines(x=c(top95freq[i], top95freq[i]), y=c(top95pc[i],ylims[2]), lty="dotted", col=color)
  points(x=top95freq[i], y=top95pc[i], col=color)
}
axis(side=4, at=mean(pgram$spec) + sd(pgram$spec) * c(.5, 1:3), labels=c(".5", "1", "2", "3"), cex.axis=0.8)
mtext(side=4, "spectrum sd ratio", line=1)
```

>>>>

```{r cleanup, echo=FALSE}
out_domain <- (seis$min < domain[1] | seis$max > domain[2])
invalid_values <- (seis$min == -1 & seis$max == 1) | (seis$min == 0 & seis$max == 0) | out_domain
"Removing " %s+% sum(out_domain) %s+% " values out of domain, " %s+% (sum(invalid_values) - sum(out_domain)) %s+% " explicit NA"
seis <-
  seis %>%
  mutate(min=replace(min, invalid_values, NA), max=replace(max, invalid_values, NA)) %>%
  drop_na()
# %>% mutate(belgian_second=with_tz(utc_second, "Europe/Brussels"))
remove(out_domain, invalid_values) # free ~350 Mb from RAM
seis.present <- length(seis$utc_second)
seis.total <- max(as.numeric(seis$utc_second)) - min(as.numeric(seis$utc_second))
seis.missing <- seis.total - seis.present
seis.present %s+% " present values, " %s+% seis.missing %s+% " missing points (" %s+% print_dec(seis.present / seis.total * 100, 2) %s+% "% / " %s+% print_dec(seis.missing / seis.total * 100, 2) %s+% "%)"
```

@@

```{r}
head(seis.h)
tides <- read.csv("dailymax.csv")
days <-seq(from=as.Date("1983-1-01"), length.out = nrow(tides), by="day")
ts.tides <- xts(tides$Level, order.by = days, frequency = 365)
#xts objects have frequency set to daily by default, so we change it
attr(ts.tides, 'frequency') <- 365
plot(stl(as.ts(ts.tides), s.window = "periodic"))
```

@@

fitted <- fitma.list[order(aic_vals$AIC)][[1]]
fitted
tsdisplay(residuals(fitted))

@@

```{r fourier}
seis.h.start <- with_tz(as_datetime('2018-11-01T00:00:00'), "Europe/Brussels")
seis.h.end <- with_tz(as_datetime('2020-03-01T00:00:00'), "Europe/Brussels")
seis.h.msts <- build_ts(seis.h, periods = c(24, 168), start = seis.h.start, end = seis.h.end)
fitted <- auto.arima(seis.h.msts, D=0, max.P=0, max.Q=0, xreg=fourier(seis.h.msts, c(3, 4)))
plot(fitted)
plot(residuals(fitted))
tsdisplay(residuals(fitted))
res.xts <- ts_zoo(residuals(fitted))
res.msts <- msts(res.xts, c(24, 168))
res.aic_vals_temp <- NULL
res.aic_vals <- NULL
for (i in 1:5) {
  for (j in 1:5) {
    #for (k in 1:5) {
      res.xreg.d <- fourier(res.msts, c(i, j)  )  # daily  seasonality
      res.fitma <- auto.arima(res.msts, D=0, max.P=0, max.Q=0, xreg=res.xreg.d)
      res.aic_vals_temp <- cbind(i, j, res.fitma$aic)  #,k
      res.aic_vals <- rbind(res.aic_vals, res.aic_vals_temp)
      print("(" %s+% i %s+% ", " %s+% j %s+% ") = " %s+% res.fitma$aic)
    #}
  }
}
remove(res.aic_vals_temp, res.xreg.d, res.fitma)
colnames(res.aic_vals) <- c("Fourier24", "Fourier168", "AIC")
res.aic_vals <- data.frame(res.aic_vals)
res.min_aic <- min(res.aic_vals$AIC)
res.minvals <- res.aic_vals[which(res.aic_vals$AIC == res.min_aic),]
res.p <- ggplot(res.aic_vals, aes(Fourier24, Fourier168)) + geom_tile(aes(fill = AIC), colour = "white") + scale_fill_gradient(low = "white", high = "steelblue")
plot(res.p)
res.minvals
res.fitted <- auto.arima(res.msts, D=0, max.P=0, max.Q=0, xreg=fourier(res.msts, c(res.minvals$Fourier24, res.minvals$Fourier168)))
plot(res.fitted)
tsdisplay(residuals(fitted))
tsdisplay(residuals(res.fitted))
```

```{r periodgram}
library(spec)
pgram <- spec.pgram(build_ts(seis.h, start=as_datetime('2018-11-01T00:00:00'), end=as_datetime('2020-03-01T00:00:00')), main="periodgram", detrend=F, demean=T, na.action=na.aggregate)
#pgram <- spec.pgram(build_ts(seis.h, start=as_datetime('2019-05-03T00:00:00'), end=as_datetime('2020-03-01T00:00:00')), main="periodgram", detrend=F, demean=T, na.action=na.aggregate)

library(spec)
pgram <- spec.pgram(build_ts(seis.h, start=as_datetime('2018-11-01T00:00:00'), end=as_datetime('2020-03-01T00:00:00')), main="periodgram", detrend=F, demean=T, na.action=na.aggregate)
topfreq <- pgram$freq[order(pgram$spec, decreasing = T)][1:100]
topseas <- 1/topfreq
topseas.h <- topseas/3600

top95pc <- pgram$spec[pgram$spec > (mean(pgram$spec) + sd(pgram$spec)*2)]
top95freq <- pgram$freq[pgram$spec > (mean(pgram$spec) + sd(pgram$spec)*2)][order(top95pc, decreasing = T)]
top95pc <- pgram$spec[pgram$spec > (mean(pgram$spec) + sd(pgram$spec)*2)][order(top95pc, decreasing = T)]
top95seas.s <- 1/top95freq
top95seas.s
top95seas.h <- top95seas.s/3600
top95pc/sd(pgram$spec)
top95seas.h

ylims <- 10^par("usr")[3:4]
axis(side=3, at=top95freq, labels=as.integer(top95seas.h), padj = 1, las=2, cex.axis=0.8)
abline(h=mean(pgram$spec) + sd(pgram$spec)*c(.3,1:3), lty="dotted", col=c("orange", "darkgreen", "blue", "red"))
for (i in 1:length(top95freq)) {
  color <- ifelse(top95pc[i] > (mean(pgram$spec) + sd(pgram$spec)*3), "red", ifelse(top95pc[i] > (mean(pgram$spec) + sd(pgram$spec)*2), "blue", ifelse(top95pc[i] > (mean(pgram$spec) + sd(pgram$spec)*1), "darkgreen", "orange")))
  lines(x=c(top95freq[i], top95freq[i]), y=c(top95pc[i],ylims[2]), lty="dotted", col=color)
  points(x=top95freq[i], y=top95pc[i], col=color)
}
axis(side=4, at=mean(pgram$spec) + sd(pgram$spec) * c(.3, 1:3), labels=c(".3", "1", "2", "3"), cex.axis=0.8)
mtext(side=4, "spectrum sd ratio", line=1)
```
